---
title: "Osa Conservation: Lifetime exploration"
author:
  - Chris Beirne
date: "`r format(Sys.Date(), '%B %d, %Y')`"
site: bookdown::test-bookdown
output: bookdown::gitbook
documentclass: book
biblio-style: apalike
link-citations: yes
editor_options: 
  chunk_output_type: console
---

# Raw data

This section summarizes the movement locations for all the tracking devices within Osa Conservation's movement ecology project. 

```{r c01, echo=F, message=F, include=F}
knitr::opts_chunk$set(echo = FALSE, message=F, warning=F)

#library(rgdal)  
library(move)
library(dplyr)
library(leaflet)
library(units)
library(sf)
library(viridis)
library(kableExtra)
library(lubridate)
library(plotly)
## Load packages for google drive ----
library(googledrive)
library(purrr)
library(readxl)
library(geosphere)
library(foreach)
#library(maptools)
library(leaflet.extras)
#library(rnaturalearth)
library(terra)
library(ggplot2)
library(atlastools)
library(ctmm)
library(move2)
 
options(googledrive_quiet = TRUE)

# For when the trapping effort file is sorted
googledrive::drive_auth(path = Sys.getenv("GOOGLE_AUTHENTICATION_CREDENTIALS"))

# ## Find Google Drive folder 'Centre Circle Data & Info'
# data_path <- "data" 
# dir.create(data_path) 

# Get the habitat raster 
#drive_download(as_id("1zDN1uh_ioOdIFATp9Q18bYTukygNo52t"), overwrite=T)

# Import passcodes
MOVE_PASS <- Sys.getenv("MOVEBANK_PASSWORD")
MOVE_USE  <- Sys.getenv("MOVEBANK_USERNAME")

loginStored <- movebankLogin(username=MOVE_USE, 
                             password=MOVE_PASS)

# Get animals
# Vultures
animals <-getMovebankAnimals(study=1573471517,login=loginStored)
# Ocelot
tmp <-getMovebankAnimals(study=2526574641,login=loginStored)
# Tapir
tmp2 <- getMovebankAnimals(study=1954804459,login=loginStored)
# Carnivores
tmp3 <- getMovebankAnimals(study=2957292408,login=loginStored)
# andean condor foundation
tmp4 <- getMovebankAnimals(study=2885579493,login=loginStored)
# Peccary
tmp5 <- getMovebankAnimals(study=		5956356622,login=loginStored)
tmp5 <- tmp5[tmp5$local_identifier!="TEST",]



# Move nickname to locak_identifier
tmp4$local_identifier <- tmp4$nick_name
tmp4$animalName <- tmp4$nick_name

animals <- rbind(animals, tmp,tmp2, tmp3, tmp5)

# They are duplicated as there are two types of data - subset to GPS
# They vary by the field "sensor_type_id"
animals <- animals[animals$sensor_type_id==653 & is.na(animals$sensor_type_id)==F,]

# Clean up the name
animals$animalName <- paste0(sub('\\_.*', '', animals$animalName), "_", sub('\\ .*', '', animals$taxon_canonical_name))
animals$name <- sub('\\_.*', '', animals$animalName)

# Sort date objects
animals$timestamp_start <- ymd_hms(animals$timestamp_start)
animals$timestamp_end <- ymd_hms(animals$timestamp_end)

t <- now("America/Costa_Rica")
#start_t <- t-as.difftime(14,units='days')
#start_mammal <- t-as.difftime(48,units='days')

# Vultures
tmp1 <- getMovebankData(study=1573471517, login=loginStored,  removeDuplicatedTimestamps=TRUE)

#OCelot - removed as collar no longer reporting
#tmp2 <- getMovebankData(study=2526574641, login=loginStored,  removeDuplicatedTimestamps=TRUE)

#Tapir
tmp3 <- getMovebankData(study=1954804459, login=loginStored,  removeDuplicatedTimestamps=TRUE)
# Remove the obvious outlier
tmp3 <- tmp3[tmp3$location_lat>8,]

#Carnivores
tmp4 <- getMovebankData(study=2957292408, login=loginStored,  removeDuplicatedTimestamps=TRUE)

#Andean condor foundation
#tmp5 <- getMovebankData(study=2885579493, login=loginStored,  #removeDuplicatedTimestamps=TRUE)

#Peccary project
tmp6 <- getMovebankData(study=5956356622, login=loginStored,  removeDuplicatedTimestamps=TRUE)


#ec_names <- data.frame(code=row.names(tmp5@idData), name=(tmp5@idData$nick_name))

mov_dat <- moveStack(tmp1, 
                     tmp3, 
                     tmp4, 
                     #tmp5,
                     tmp6)
# Remove the random zeros
summary(mov_dat$location_long)
mov_dat <- mov_dat[mov_dat$location_long<0,]





#Add the names
mov_dat$name <- trackId(mov_dat)


# Convert timezone
mov_dat$timestamp <- with_tz(timestamps(mov_dat), tz="America/Costa_Rica")
 
# all data
#dat <- getMovebankData(study=1573471517, login=loginStored,  removeDuplicatedTimestamps=TRUE)

# Convery move stack to dataframe
dat <- as.data.frame(mov_dat)

# Convert dat to costa rica time
dat$timestamp <- with_tz(dat$timestamp, tzone = "America/Costa_Rica")


# Add the location data
dat <- left_join(dat, animals[, c("deployment_id", "animalName")])
# Sort the names out
dat$animalName <- sub('\\_.*', '', dat$animalName)

# Add in the taxonomic group
dat$animalName <- paste0(dat$animalName, "_", sub('\\ .*', '', dat$taxon_canonical_name))

# # Setup the leaflet icon
# leafIcons <- icons(
#   iconUrl = "https://raw.githubusercontent.com/ChrisBeirne/Osa-Conservation-Movement-Ecology-Daily-Report/main/icons/vul_small1.png",
#   iconWidth = 38, iconHeight = 40,
#   iconAnchorX = 22, iconAnchorY = 39,
#   shadowUrl = "https://raw.githubusercontent.com/ChrisBeirne/Osa-Conservation-Movement-Ecology-Daily-Report/main/icons/vul_small_shadow.png",
#   shadowWidth = 38, shadowHeight = 30,
#   shadowAnchorX = 4, shadowAnchorY = 39
# )
# #?iconList()

papa <- makeIcon(
  iconUrl = "https://raw.githubusercontent.com/ChrisBeirne/Osa-Conservation-Movement-Ecology-Daily-Report/main/icons/king_small.png",
  iconWidth = 38, iconHeight = 40,
  iconAnchorX = 22, iconAnchorY = 39,
  shadowUrl = "https://raw.githubusercontent.com/ChrisBeirne/Osa-Conservation-Movement-Ecology-Daily-Report/main/icons/vul_small_shadow.png",
  shadowWidth = 38, shadowHeight = 30,
  shadowAnchorX = 4, shadowAnchorY = 39)

aura <- makeIcon(
  iconUrl = "https://raw.githubusercontent.com/ChrisBeirne/Osa-Conservation-Movement-Ecology-Daily-Report/main/icons/turk_small.png",
  iconWidth = 38, iconHeight = 40,
  iconAnchorX = 22, iconAnchorY = 39,
  shadowUrl = "https://raw.githubusercontent.com/ChrisBeirne/Osa-Conservation-Movement-Ecology-Daily-Report/main/icons/vul_small_shadow.png",
  shadowWidth = 38, shadowHeight = 30,
  shadowAnchorX = 4, shadowAnchorY = 39)

mela <- makeIcon(
  iconUrl = "https://raw.githubusercontent.com/ChrisBeirne/Osa-Conservation-Movement-Ecology-Daily-Report/main/icons/yhv_small.png",
  iconWidth = 38, iconHeight = 40,
  iconAnchorX = 22, iconAnchorY = 39,
  shadowUrl = "https://raw.githubusercontent.com/ChrisBeirne/Osa-Conservation-Movement-Ecology-Daily-Report/main/icons/vul_small_shadow.png",
  shadowWidth = 38, shadowHeight = 30,
  shadowAnchorX = 4, shadowAnchorY = 39)

pardalis <- makeIcon(
  iconUrl = "https://raw.githubusercontent.com/ChrisBeirne/Osa-Conservation-Movement-Ecology-Daily-Report/main/icons/ocelot.png",
  iconWidth = 38, iconHeight = 40,
  iconAnchorX = 22, iconAnchorY = 39)

bairdii <- makeIcon(
  iconUrl = "https://raw.githubusercontent.com/ChrisBeirne/Osa-Conservation-Movement-Ecology-Daily-Report/main/icons/tapir.png",
  iconWidth = 38, iconHeight = 40,
  iconAnchorX = 22, iconAnchorY = 39)

microtis <- makeIcon(
  iconUrl = "https://raw.githubusercontent.com/ChrisBeirne/Osa-Conservation-Movement-Ecology-Daily-Report/main/icons/sed.png",
  iconWidth = 38, iconHeight = 40,
  iconAnchorX = 22, iconAnchorY = 39)

atra <- makeIcon(
  iconUrl = "https://raw.githubusercontent.com/ChrisBeirne/Osa-Conservation-Movement-Ecology-Daily-Report/main/icons/bhv_small.png",
  iconWidth = 38, iconHeight = 40,
  iconAnchorX = 22, iconAnchorY = 39,
  shadowUrl = "https://raw.githubusercontent.com/ChrisBeirne/Osa-Conservation-Movement-Ecology-Daily-Report/main/icons/vul_small_shadow.png",
  shadowWidth = 38, shadowHeight = 30,
  shadowAnchorX = 4, shadowAnchorY = 39)

onca <- makeIcon(
  iconUrl = "https://raw.githubusercontent.com/ChrisBeirne/Osa-Conservation-Movement-Ecology-Daily-Report/main/icons/jaguar.png",
  iconWidth = 38, iconHeight = 36,
  iconAnchorX = 22, iconAnchorY = 39)

pecari <- makeIcon(
  iconUrl = "https://raw.githubusercontent.com/ChrisBeirne/Osa-Conservation-Movement-Ecology-Daily-Report/main/icons/WLP.png",
  iconWidth = 38, iconHeight = 30,
  iconAnchorX = 22, iconAnchorY = 39)

iconSet <- iconList(aura= aura,
                    papa =papa,
                    melambrotus = mela,
                    atratus = atra,
                    pardalis = pardalis,
                    bairdii = bairdii,
                    microtis = microtis,
                    onca=onca,
                    pecari=pecari)


dat$icon <- sub(".*? ", "", dat$taxon_canonical_name)


# Add costa rica survey locations
cr_locs<- data.frame(location_code=c("AIRS02", "CERR02","LOMAS01"), 	
           latitude=c(8.402595,	8.416142,8.758321)	,
           longitude=c( -83.35852,	-83.31793,	-83.29787) )

# Add peru survey locations
pe_locs<- data.frame(location_code=c("VT_MANU01"), 	
           latitude=c(-12.887758)	,
           longitude=c( -71.403264) )


```

*Last data request: `r now(tzone = "America/Costa_Rica")`*

The project currently contains **`r length(unique(dat$tag_id))` tag deployments** and **`r nrow(dat)` locations**. 

```{r  c02, include=F}
# Create data summaries 

# Remove the crazy ocelot location in ecuador
dat <- dat[!(dat$location_lat>(-6) & dat$location_lat<6 & dat$taxon_canonical_name=="Leopardus pardalis"),]
dat <- dat[!(dat$location_lat>(-2) & dat$location_lat<(-1) & dat$taxon_canonical_name=="Coragyps atratus"),]

# Add a country column
dat$country <- NA
dat$country[dat$location_lat>6] <- "costa_rica"
dat$country[dat$location_lat>(-6) & dat$location_lat<6] <- "ecuador"
dat$country[dat$location_lat<(-6)] <- "peru"

#table(dat$country)

# Costa rica
# Convert to shapefiles
dat <- dat[order(dat$animalName),]


# Rename the ecuador tags
dat$name<- as.character(dat$name)
i <- 1
for(i in 1: nrow(ec_names))
  {
    dat$name[dat$name==ec_names$code[i]] <- ec_names$name[i]      
  }
dat$name<- as.factor(dat$name)



lfc <- do.call(st_sfc,
              lapply(split(dat, dat$name),
                     function(d){st_linestring(as.matrix(d[,c("location_long", "location_lat")]))}))
dat_shp <- st_sf(data.frame(name=levels(factor(dat[,"name"])), geom=lfc))

#plot(dat$location_long[dat$country=="costa_rica"], dat$location_lat[dat$country=="costa_rica"])
#plot(st_geometry(dat_shp), add=T)

sp_dat <- dat[,c("name",    "country", "taxon_canonical_name" )]
sp_dat <- sp_dat[duplicated(sp_dat)==F,]

# add back in the metadata
dat_shp  <- left_join(dat_shp,sp_dat)

# Last location and time since present
dat$diff_time <- round(as.numeric(difftime(t, dat$timestamp, units="days")),1)

last_obs <- dat %>% group_by(name) %>% dplyr::summarize(days_since_fix=min(diff_time))

last_sum  <- dat %>% 
  group_by(name) %>%
  slice(which.max(timestamp))


tmp <- last_sum[, c("name","taxon_canonical_name","country", "timestamp", "diff_time")]
colnames(tmp)[colnames(tmp)=="diff_time"] <- "days_since_checkin"
colnames(tmp)[colnames(tmp)=="timestamp"] <- "last_timestamp"
check_in_summary <- tmp[order(tmp$country, tmp$days_since_checkin),]


# For chris - just vultures
# trans <- dat[!(dat$height_above_ellipsoid>5000 & is.na(dat$height_above_ellipsoid)==F) ,]
# trans <- trans[trans$taxon_canonical_name %in% c("Cathartes aura", "Cathartes melambrotus",      "Coragyps atratus" ,"Sarcoramphus papa") ,]
# trans$name <- factor(trans$name)
# lfc <- do.call(st_sfc,
#               lapply(split(trans, trans$name),
#                      function(d){st_linestring(as.matrix(d[,c("location_long", "location_lat")]))}))
# trans_shp <- st_sf(data.frame(name=levels(factor(trans[,"name"])), geom=lfc))
# st_write(trans_shp,"C:/Users/cwbei/Dropbox/Osa Conservation/Data analysis/current_vulture_data/all_vulture_tracks.shp", append=FALSE )
# write.csv(trans, "C:/Users/cwbei/Dropbox/Osa Conservation/Data analysis/current_vulture_data/all_vulture_points.csv", row.names=F)
#table(dat$taxon_canonical_name, dat$country)

```

The number of species by country is:

```{r c03, echo=F}

data <- t(as.data.frame.matrix(table(check_in_summary$taxon_canonical_name, check_in_summary$country)))
data <- as.data.frame(data)
data$country <- row.names(data)
fig <- plot_ly(data, x = ~country, y = ~data$`Sarcoramphus papa`, type = 'bar', name = 'King vulture')
fig <- fig %>% add_trace(y = ~data$`Cathartes melambrotus`, name = 'GYH vulture')
fig <- fig %>% add_trace(y = ~data$`Cathartes aura`, name = 'Turkey vulture')
fig <- fig %>% add_trace(y = ~data$`Coragyps atratus`, name = 'Black vulture')
fig <- fig %>% add_trace(y = ~data$`Tapirus bairdii`, name = 'Tapir')
fig <- fig %>% add_trace(y = ~data$`Panthera onca`, name = 'Jaguar')
fig <- fig %>% add_trace(y = ~data$`Puma concolor`, name = 'Puma')
fig <- fig %>% add_trace(y = ~data$`Atelocynus microtis`, name = 'SE dog')
fig <- fig %>% add_trace(y = ~data$`Leopardus pardalis`, name = 'Ocelot')
fig <- fig %>% add_trace(y = ~data$`Tayassu pecari`, name = 'WLP')
fig <- fig %>% layout(yaxis = list(title = 'Count'), barmode = 'stack')

fig
```

Locations per country and species:

```{r c03a, echo=F}
tmp <- dat %>% group_by(country ,taxon_canonical_name) %>% summarise(n_locs=n(), n_inds=n_distinct(local_identifier)) 

tmp %>%
  kbl(caption = "Summary: Locations and individuals by country") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

Temporal patterns in species tag deployments:

*note* if an animal is recaptured and a new device is installed this will count as two deployments. 

```{r}
# head(animals)
# 
# table(animals$taxon_canonical_name)

i <- 1
time_sum <- data.frame(year=unique(substr(animals$timestamp_start,1,4)), total_tags_deployed=NA, mammal_tags=NA, vulture_tags=NA)

# remove NA's
time_sum <- time_sum[is.na(time_sum$year)==F,]


for(i in 1:nrow(time_sum))
{
  tmp <- animals[substr(animals$timestamp_start,1,4)==time_sum$year[i],]
  time_sum$total_tags_deployed[i] <- nrow(tmp)
  time_sum$mammal_tags[i] <- nrow(tmp[!(tmp$taxon_canonical_name %in% c("Cathartes melambrotus", "Sarcoramphus papa","Cathartes aura", "Coragyps atratus")),])
  time_sum$vulture_tags[i] <- nrow(tmp[tmp$taxon_canonical_name %in% c("Cathartes melambrotus", "Sarcoramphus papa","Cathartes aura", "Coragyps atratus"),])
}

time_sum %>%
  kbl(caption = "Summary: tag deployments through time") %>%
  kable_classic(full_width = F, html_font = "Cambria")


```


### Costa Rica

**Note this is raw unfiltered data**

```{r c04, echo=F}
tmp_cr <- dat[dat$country=="costa_rica",]
tmp_cr_shp <- dat_shp[dat_shp$country=="costa_rica",]

# First lets choose a category to colour
tmp_cr[,"name"] <- factor(tmp_cr[,"name"])
tmp_cr_shp$name <- factor(tmp_cr_shp$name)

col.cat <- turbo(length(levels(tmp_cr[,"name"])))
# Add it to the dataframe
tmp_cr$colours <- col.cat[tmp_cr[,"name"]]
tmp_cr_shp$colours <- col.cat[tmp_cr_shp$name]

```

We currently have deployed *`r length(unique(tmp_cr$name))`* devices in Costa Rica:

```{r c05, echo=F}

# New map
ids <- tmp_cr_shp$name
names <- sub('\\_.*', '', tmp_cr_shp$name)
i <- 5
m <- leaflet() %>%
  # Add a satellite image layer
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
  addProviderTiles(providers$OpenStreetMap.Mapnik, group="OS") %>%  
    addProviderTiles(providers$CartoDB.PositronNoLabels, group="Simple")   

  for(i in seq(along=ids))
    {
      m <- m %>%
        addPolylines(data = tmp_cr_shp[i,], color = tmp_cr_shp$colours[i], group = ids[i], weight=1,opacity=0.5, popup = ids[i])
        
        
    }

  m <- m %>%
      # Add survey location
      addMarkers(lng=cr_locs$longitude,
                         lat=cr_locs$latitude, 
                         popup=cr_locs$location_code) %>% 
      #addLegend(position= "topright", colors=tmp_cr_shp$colours, 
       #         labels=names ,opacity = 0.7, title = "Animals") %>%
      addScaleBar(position="topleft", 
                  options=scaleBarOptions(maxWidth = 100, metric = TRUE, imperial = FALSE, updateWhenIdle = TRUE)) %>%
      addLayersControl(
        baseGroups = c("Simple", "OS", "Satellite"),
        overlayGroups = ids,
        options = layersControlOptions(collapsed = FALSE)
      ) %>%
  addFullscreenControl()
  
m 
```


### Peru

```{r c06, echo=F}
tmp_peru <- dat[dat$country=="peru",]
tmp_peru_shp <- dat_shp[dat_shp$country=="peru",]


# First lets choose a category to colour
tmp_peru[,"name"] <- factor(tmp_peru[,"name"])
tmp_peru_shp$name <- factor(tmp_peru_shp$name)

col.cat <- turbo(length(levels(tmp_peru[,"name"])))
# Add it to the dataframe
tmp_peru$colours <- col.cat[tmp_peru[,"name"]]
tmp_peru_shp$colours <- col.cat[tmp_peru_shp$name]

```

We have deployed `r length(unique(tmp_peru$name))` devices in Peru:


```{r c07, echo=F}
# New map
ids <- tmp_peru_shp$name
i <- 1
m <- leaflet() %>%
  # Add a satellite image layer
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
  addProviderTiles(providers$OpenStreetMap.Mapnik, group="OS") %>%  
    addProviderTiles(providers$CartoDB.PositronNoLabels, group="Simple")


  for (i in seq(along=ids))
    {
      m <- m %>%
        addPolylines(data = tmp_peru_shp[i,], color = tmp_peru_shp$colours[i], group = ids[i], weight=1,opacity=0.5)  
    }
    
  m <- m %>%
      addScaleBar(position="topleft", 
                  options=scaleBarOptions(maxWidth = 100, metric = TRUE, imperial = FALSE, updateWhenIdle = TRUE)) %>%
      addLayersControl(
        baseGroups = c("Simple","OS", "Satellite"),
        overlayGroups = ids,
        options = layersControlOptions(collapsed = FALSE)
      ) %>%
  addFullscreenControl()
  
m 
```

### Ecuador

```{r c08, echo=F}
tmp_ecuador <- dat[dat$country=="ecuador",]
tmp_ecuador_shp <- dat_shp[dat_shp$country=="ecuador",]


# First lets choose a category to colour
tmp_ecuador[,"name"] <- factor(tmp_ecuador[,"name"])
tmp_ecuador_shp$name <- factor(tmp_ecuador_shp$name)

col.cat <- turbo(length(levels(tmp_ecuador[,"name"])))
# Add it to the dataframe
tmp_ecuador$colours <- col.cat[tmp_ecuador[,"name"]]
tmp_ecuador_shp$colours <- col.cat[tmp_ecuador_shp$name]

```

We have deployed `r length(unique(tmp_ecuador$name))` devices in Ecuador:


```{r c09, echo=F}
# New map
ids <- tmp_ecuador_shp$name
i <- 1
m <- leaflet() %>%
  # Add a satellite image layer
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
  addProviderTiles(providers$OpenStreetMap.Mapnik, group="OS") %>%  
    addProviderTiles(providers$CartoDB.PositronNoLabels, group="Simple")


  for (i in seq(along=ids))
    {
      m <- m %>%
        addPolylines(data = tmp_ecuador_shp[i,], color = tmp_ecuador_shp$colours[i], group = ids[i], weight=1,opacity=0.5)  
    }
    
  m <- m %>%
      addScaleBar(position="topleft", 
                  options=scaleBarOptions(maxWidth = 100, metric = TRUE, imperial = FALSE, updateWhenIdle = TRUE)) %>%
      addLayersControl(
        baseGroups = c("Simple","OS", "Satellite"),
        overlayGroups = ids,
        options = layersControlOptions(collapsed = FALSE)
      ) %>%
  addFullscreenControl()
  
m 
```


